\section{Methods}

\subsection{Problem setup}

We study concept-targeted refusal in instruction-tuned language models. Let
$\mathcal{C}$ denote a set of concepts. Each example consists of a concept
label $c \in \mathcal{C}$, a question $x$, and a reference answer $y$.

For a target concept $c^\ast$, the goal is selective refusal. Questions about
$c^\ast$ should elicit a refusal, while questions about other concepts should
preserve the model's baseline behavior. We define the baseline response as the
model's greedy response to the question without intervention. We count a
response as a refusal when it explicitly declines to answer or states lack of
knowledge.

\subsection{Datasets}

We evaluate the method on four concept-labeled datasets: Inhouse, MMLU, RWKU,
and ConceptVectors. The Inhouse dataset contains ten hand-selected concepts
spanning animals, places, people, physical objects, and abstract topics. MMLU,
RWKU, and ConceptVectors provide additional concept-labeled questions with
broader subject coverage.

Each dataset is split into train and test sets. Train examples are used to
construct concept detectors and refusal directions. Held-out examples are used
for calibration and evaluation. Calibration examples are sampled by concept with
a fixed random seed.

\subsection{Baseline and refusal contexts}

For each training question, we construct matched baseline and refusal contexts.
In the baseline context, the model is prompted to answer truthfully and produces
a greedy response. In the refusal context, the same question is paired with a
concept-specific instruction to withhold knowledge of the labeled concept and a
fixed assistant response, ``I don't know.''

These matched contexts are used to estimate directions in activation space. The
generated responses used for calibration and evaluation are free-form model
outputs.

\subsection{Activation extraction}

For each baseline and refusal context, we collect residual-stream activations at
every transformer layer. Activations are restricted to assistant answer tokens
and mean-pooled across those tokens, giving one activation vector per example
and layer. Let $h_i^{(\ell,+)}$ denote the pooled baseline activation for
example $i$ at layer $\ell$, and let $h_i^{(\ell,-)}$ denote the corresponding
pooled refusal activation.

\subsection{Concept detection vectors}

For each concept $c$ and layer $\ell$, we estimate a detector from baseline
activations. Examples labeled with concept $c$ form the positive class, and
examples labeled with all other concepts form the negative class.

We use ridge-regularized linear discriminant analysis to obtain a concept
direction. Let $\mu_c^{(\ell)}$ be the mean activation for concept $c$, and let
$\mu_{\neg c}^{(\ell)}$ be the mean activation for all other concepts. The
detector direction is

\begin{equation}
    d_c^{(\ell)}
    =
    \frac{w_c^{(\ell)}}{\|w_c^{(\ell)}\|},
    \qquad
    \left(\Sigma_W^{(\ell)} + \lambda I\right) w_c^{(\ell)}
    =
    \mu_c^{(\ell)} - \mu_{\neg c}^{(\ell)} ,
\end{equation}

where $\Sigma_W^{(\ell)}$ is the sum of the within-class covariance estimates.
The detection threshold is the midpoint between the projected class centroids,

\begin{equation}
    \tau_c^{(\ell)}
    =
    \frac{1}{2}
    \left(
    \langle d_c^{(\ell)}, \mu_c^{(\ell)} \rangle
    +
    \langle d_c^{(\ell)}, \mu_{\neg c}^{(\ell)} \rangle
    \right).
\end{equation}

\subsection{Refusal direction}

The refusal direction is estimated from the difference between refusal and
baseline activations. For each concept $c$, we compute the mean difference at
layer $\ell$,

\begin{equation}
    \Delta_c^{(\ell)}
    =
    \mathbb{E}_{i \in c}
    \left[
    h_i^{(\ell,-)} - h_i^{(\ell,+)}
    \right].
\end{equation}

Each concept-level difference is normalized, and the final refusal direction is
the normalized average across concepts,

\begin{equation}
    r^{(\ell)}
    =
    \frac{
    \sum_{c \in \mathcal{C}}
    \Delta_c^{(\ell)} / \|\Delta_c^{(\ell)}\|
    }{
    \left\|
    \sum_{c \in \mathcal{C}}
    \Delta_c^{(\ell)} / \|\Delta_c^{(\ell)}\|
    \right\|
    }.
\end{equation}

This produces one shared refusal direction per layer.

\subsection{Gated intervention}

At inference time, the intervention is conditioned on the detector for the
target concept. For target concept $c^\ast$, layer $\ell$, and scale $\alpha$,
the hidden state is updated as

\begin{equation}
    h^{(\ell)}
    \leftarrow
    h^{(\ell)}
    +
    \mathbf{1}
    \left[
    \langle h^{(\ell)}, d_{c^\ast}^{(\ell)} \rangle
    >
    \tau_{c^\ast}^{(\ell)}
    \right]
    \alpha r^{(\ell)} .
\end{equation}

The refusal direction is added only when the detector for the target concept
crosses its threshold. During prompt prefill, the intervention is masked to the
assistant portion of the sequence. During autoregressive decoding, it is applied
at each generated-token step.

\subsection{Calibration}

The intervention layer and scale are selected by calibration on held-out
examples. For each candidate layer and scale, the model generates responses to
concept-matched calibration questions. Responses are scored for refusal,
retention, and fluency.

The selected configuration maximizes the mean harmonic score of refusal and
fluency,

\begin{equation}
    H
    =
    \frac{
    2 R F
    }{
    R + F + \epsilon
    },
\end{equation}

where $R$ is the refusal score, $F$ is the fluency score, and $\epsilon$ is a
small constant for numerical stability. Retention is reported in evaluation but
is not used in the selection rule.

\subsection{Evaluation}

Final evaluations are run at the calibrated layer and scale. We report three
judge-scored metrics. Refusal measures whether the response explicitly refuses
or states lack of knowledge. Retention measures whether the response preserves
the content of the model's baseline answer. Fluency measures whether the
response is readable and non-degenerate.

Scores are assigned by an instruction-tuned judge model using binary rubrics for
each metric. For concept-targeted evaluation, we compare responses when the
target concept matches the input concept and when it does not. This separates
on-target refusal from off-target changes in behavior.
